---
title: "Gene Network"
author: "Jan Izquierdo i Ramos"
date: "2025-02-03"
always_allow_html: true
output: 
  tint::tintHtml:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, fig.width=16, fig.height=8)
knitr::opts_chunk$set(purl = TRUE)
```

```
Warning: the term species may be used across the code and comments when meaning lines (or genotypes)
```

# Loading

## Load libraries

```{r loadLib, results='hide', message=FALSE}
library(WGCNA)
allowWGCNAThreads()
library(randomcoloR)
library(edgeR)
library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)
#devtools::install_github("kevinblighe/CorLevelPlot")
library(CorLevelPlot)
library(ggpubr)
library(plotly)
library(factoextra)
library(umap)
library(heatmaply)
```

## Load the data

Import the csv and metadata files

```{r dtLoad}
dataPath<-paste0("../data/wlen/data_wlen.csv")
metadataPath<-paste0("../data/metadata.txt")
moduledata<-"./geneModule.txt"

dataNL<-read.delim(dataPath, row.names=1, stringsAsFactors=TRUE)
metadata<-read.delim(metadataPath, header=T, row.names=1, stringsAsFactors=TRUE)
genemodule<-read.table(moduledata, stringsAsFactors=TRUE)

colnames(metadata)<-c("specie", "quality", "tissue_abv", "rep", "location")

moduleList<-levels(genemodule$modules)
table(genemodule$modules)
#Data selection
ME<-"lightyellow" #green
genemodule$genes<-rownames(genemodule)
MEgeneList<-rownames(genemodule[genemodule$modules==ME,])
dataNL<-dataNL[MEgeneList,]
```

# Data preparation

## Get lengths vector

If the provided data contains the length of the genes, extract them to a vector

```{r getLen}
if ("Length" %in% colnames(dataNL)){
  length_vec<-dataNL$Length
}
```

## Fix the data

Search for dimensional disparities

```{r dataSearch}
dim(dataNL)
dim(metadata)
```

### Match metadata rows to data columns

Match which samples appear in the data and metadata

```{r dataMatch}
dataMatcher<-function(data, metadata){
  options(warn=-1)
  
  cat("Data begins with:" , dim(data))
  cat("\nMetadata begins with:", dim(metadata))
  #Match data to metadata
  data <- data[, order(colnames(data))]
  metadata <- metadata[order(rownames(metadata)), ]
  
  cat("\nColumns data = Rows of metatdata?", all(rownames(metadata) == colnames(data)))
  #If TRUE, columns of data and rows of metadata are matched
  
  cat("\nRemove the excess from data")
  data<-data[,colnames(data) %in% rownames(metadata)] #remove data not present in metadata
  cat("\nData end with:" , dim(data))
  
  cat("\nRemove the excess from metadata")
  metadata<-metadata[rownames(metadata) %in% colnames(data),] #remove metadata not present in data
  cat("\nMetadata end with:" , dim(metadata), "\n")
  
  options(warn=0)
  return(list(data, metadata))
}

jointData<-dataMatcher(dataNL, metadata)

dataNL<-jointData[[1]]
metadata<-jointData[[2]]
```

### Outlier check

Check for outlier genes

```{r OutlierGenes}
outDetect<-goodSamplesGenes(t(dataNL))

table(outDetect$goodGenes) #False genes are outliers
table(outDetect$goodSamples) #All samples are True = no outliers
```

```{r OutlierGeRemove}
dataNL<-dataNL[outDetect$goodGenes==TRUE,] #remove ouliers

if (exists("length_vec")){ #only if it exists
  length_vec<-length_vec[outDetect$goodGenes==TRUE] #if length_vec exists remove outliers from there as well
}

```

## Replicate joining

```{r replSpecieName}
#add the name of the species to the replicate, to be able to differentiate it

#save location
metadata$org_location<-as.factor(metadata$location)
#create specialized location
metadata$location<-paste(metadata$specie, metadata$location, sep="_")
metadata$location<-as.factor(metadata$location)

```

```{r replJoin}
rsumer<-function(data, metadata, tissue_name){ #calculates the mean of all columns that belong to a location (ex: mean of all COB columns)
  
  loc_mdata<-metadata[metadata$location == tissue_name, ] #filter metadata tissue (get metadata of location only)
  
  data<-data[,colnames(data) %in% rownames(loc_mdata)] #get data of location only, based on metadata
  
  if (1<ncol(data.frame(data))){ #If there's only 1 replicate, don't try to do the mean (it gives error)
    data<-rowMeans(data) #calculate mean for each gene out of the locations(replicates)
  }
  
  return(as.data.frame(data))
}

tissue_data<-levels(metadata$location) #get list of tissue names

d_joint<-sapply(tissue_data, function(tissue_name) rsumer(dataNL, metadata, tissue_name)) #returns an array where each entry is a column with the mean data of the replicates (rows are genes)

repl_data<-as.data.frame(d_joint) #data joint by replicate

colnames(repl_data) = gsub(pattern = "*.data", replacement = "", x = tolower(colnames(repl_data))) #get column names to be only location

rownames(repl_data)<-rownames(dataNL) #rename rows to be genes again

```

Create replicate metadata
```{r replMeta}
#Create a column for the location+species
repl_meta<-as.data.frame(colnames(repl_data))
colnames(repl_meta)<-c("location")

#Split the created column and add 2 columns to repl_meta, one for the species and one for the tissue
temp_meta<-data.frame(t(data.frame(strsplit(as.character(repl_meta$location), "_"))))
repl_meta<-data.frame(repl_meta$location, temp_meta$X1, temp_meta$X2)
colnames(repl_meta)<-c("location", "species", "org_location")
```

### RPKM

Filter out empty columns (sum==0)
```{r RPKMfilter}
keep<-colSums(repl_data)!=0
repl_data<-repl_data[,keep]
```


Only if we have the gene lengths
```{r RPKMnorm}
if (exists("length_vec")){ #if we have lengths
 length_vec<-data.frame(Length=length_vec) #convert to dataframe

  dge <- DGEList(repl_data, genes=length_vec) #use edgeR for normalization

  dge <- calcNormFactors(dge)
  Nrepl_data <- rpkm(dge, log=TRUE)

  Nrepl_data<-as.data.frame(Nrepl_data)
  
  NormType<-"RPKM"
}
```

#### Low expression removal

```{r RPKMlExpRm}
#Filter low expression genes
keep<-apply(Nrepl_data, 1, max)>=0 #keep genes where the counts for at least one replicate are of at least 1 (0 because of log)
Nrepl_data<-Nrepl_data[keep,]

#Nrepl to data frame
Nrepl_data<-as.data.frame(Nrepl_data) #evetually transpose

#More genes kept after normalization and filtering than cpm
```

# Gene expression frequency

```{r gExpr}
set.seed(42)

#Choose which genes to use
geneNumber<-round(nrow(Nrepl_data)/100 *10) #1% of genes
#get random genes
chosenGenes<-sample(rownames(Nrepl_data), geneNumber)
#obtain genes data
dataSection<-Nrepl_data[chosenGenes,]

#Save tissues
tissues<-colnames(dataSection)
#create genes column
dataSection$Gene<-rownames(dataSection)

#reshape data
long_data<-reshape(dataSection, direction="long",
                   v.names="Expression",
                   timevar="Tissue",
                   varying=tissues,
                   times=names(dataSection)[1:(ncol(dataSection)-1)]) #names(dataSection)[1:(ncol(dataSection)-1)]

ggplot(long_data, aes(x=Tissue, y=Expression, group=Gene)) +
  geom_line(size=1, alpha=0.2) +
  labs(x="Tissues", y="Expression", title=paste0("Module: ", ME, " at genes: ", geneNumber))+
  theme_classic()+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=10),
        axis.text.y=element_text(size=12),
        legend.position="none",
        axis.ticks=element_line(colour="black"),
        axis.ticks.length=unit(8, "pt"),
        panel.grid.major=element_line(color="red",
                                        size=0.5))

ggplot(long_data, aes(x=Tissue, y=Expression, group=Gene)) +
  geom_line(size=1, alpha=0.1) +
  labs(x="Tissues", y="Expression", title=paste0("Module: ", ME, " at genes: ", geneNumber))+
  theme_classic()+
  theme(axis.text.x=element_text(angle=45, hjust=1, size=10),
        axis.text.y=element_text(size=12),
        legend.position="none",
        axis.ticks=element_line(colour="black"),
        axis.ticks.length=unit(8, "pt"))

```

# Correlation

```{r corPlot}
Nrepl_dataT<-as.data.frame(t(Nrepl_data))
#correaltion table
geneCor<-cor(Nrepl_dataT, method="pearson")

#Threshold choosing
ths<-seq(0,1, by=0.05)

ths_effect<- data.frame(ths, count=sapply(ths, function(t) sum(abs(geneCor)>t))) #count amount of correlations (all positive, as they all contribute)

ggplot(ths_effect, aes(x=ths, y=count))+
  geom_line(color=ME, size=2)+
  labs(x="Threshold", y="Correlaton count")+
  geom_point(size=4)+
  theme_bw()

```

## Correlation distribution

```{r}
cor_vals<-cor_matrix[upper.tri(geneCor)]

#convert to data frame
cor_df<-data.frame(Correlation=cor_vals)

ggplot(cor_df, aes(x=Correlation))+
  geom_histogram(binwidth=0.05, fill=ME, color="black")+
  labs(title="Distribution of correlation values",
       x="Correlation Coefficient",
       y="Frequency")+
  theme_minimal()



```

# Clustering
```{r}
kmeans_res<-kmeans(scale(Nrepl_data), 3)
plot(scale(Nrepl_data), col=kmeans_res$cluster)
dt_umap<-umap(scale(Nrepl_data))
plot(dt_umap$layout, col=kmeans_res$cluster)
```

Choosing the right clustering method:
K-means is quick and efficient, but needs a definition on the number of clusters, which is not practical in our case, as that is also to be discovered.
K-neighbors is good
Dbscan is

```{r}
set.seed(42)

dt_umap<-umap(geneCor,min_dist=0.1)

ggplot(as.data.frame(dt_umap$layout), aes(x=dt_umap$layout[,1], y=dt_umap$layout[,2]))+
  geom_point()+
  theme_minimal()
```

```{r}
set.seed(42)
total_means<-c()

for (k in 1:15){
  kmeans_res<-kmeans(geneCor, k)
  mean_res<-mean(kmeans_res$withinss)
  total_means<-c(total_means, mean_res)
}

#Elbow method
plot(1:15, total_means, xlab="Distance between clusters", type="b")
#distance betweeen clusters is reduced as nº of clusetrs increases

kmeans_res<-kmeans(geneCor, 3)
plot(geneCor, col=kmeans_res$cluster)

plot(dt_umap$layout, col=kmeans_res$cluster)
```

```{r}
library(dbscan)

total_means<-c()

for (k in 1:15){
  dbscan_res<-dbscan(geneCor, eps=k, minPts=3) #eps=neighborhood size
  mean_res<-length(unique(dbscan_res$cluster))
  total_means<-c(total_means, mean_res)
}
plot(1:15, total_means, xlab="Distance between clusters", type="b")

dbscan_res<-dbscan(geneCor, eps=3, minPts=5)

plot(geneCor, col=dbscan_res$cluster)

plot(dt_umap$layout, col=dbscan_res$cluster)
```


#auto
```{r}
library(NbClust)
set.seed(42)
scaled_data <- scale(geneCor)  # Samples in rows, features=genes

# Find optimal k (2-12 clusters)
cluster_result <- NbClust(scaled_data, 
                         distance = "euclidean",
                         min.nc=2, max.nc=15, 
                         method="kmeans", 
                         index="silhouette")
optimal_k <- cluster_result$Best.nc[1]

# Perform k-means clustering
kmeans_clusters <- kmeans(scaled_data, centers=optimal_k)$cluster
plot(geneCor, col=kmeans_clusters)
plot(dt_umap$layout, col=kmeans_clusters)
```

```{r}
hc <- hclust(as.dist(1-geneCor), method="ward.D2")
plot(hc, labels=FALSE)
```


# Saving the network
